Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-13 paper-2024-sac-marl-neighbouring-strategies.tex
final pass: complete

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{figures/behavior.pdf}
  \caption{Examples of cooperative and uncoordinated behavior. (Left) desired, agents effectively choose different paths to collect different items. (Right) undesired, agents cover the same path to collect the same item.}
  \label{fig:behavior}
\end{figure}

@@

Regarding the proposed neighbor-based approaches, it can be observed that both \emph{experience sharing} and 
 \emph{NN averaging} enhance the performance of DTDE, achieving results comparable to CTDE while also increasing 
 learning stability. 
% 
On the other hand, \emph{NN consensus} maintains performance levels similar to DTDE, which may be attributed to the 
 simplicity of the implemented consensus policy. 
%
Nevertheless, it remains more stable and achieves better performance than MAPPO.

\paragraph{Scalability}
To evaluate the scalability of the learned policies, we conducted several experiments by increasing the number 
 of agents -- specifically to 8, 16, and 32 -- and measuring the time required to complete the task 
 (i.e., mean episode length).

The results -- \Cref{fig:time-eval} -- show that as the number of agents increases, DTDE exhibits both a higher mean time 
 to solve the task and greater instability. 
%
In contrast, CTDE, NN averaging, and experience sharing maintain strong performance. 
%
Similar to the training phase, NN consensus does not achieve the same results as the other two neighbor-based 
 approaches but remains on par with DTDE in terms of performance.

\paragraph{Communication overhead}

The proposed neighbor-based approaches, particularly NN averaging and experience sharing, achieve performance
 comparable to CTDE while maintaining a decentralized learning structure similar to DTDE. 
% 
However, this entails a communication overhead due to the information agents must exchange with their neighbors
 -- see \Cref{fig:scalability}. 
%
For NN consensus, the overhead remains constant, as each agent selects the best neighbor and only exchanges 
 information with that agent, meaning the amount of data shared does not increase with the number of neighbors. 
% 
In contrast, for NN averaging and experience sharing, the overhead scales linearly. 
%
Specifically, the growth in overhead for experience sharing is minimal, as the agents' experiences require very 
 little memory. 
%
For NN averaging, however, this increase is more significant, as even a moderately complex neural network 
 has a substantially larger memory footprint.
% 
Nevertheless, for all approaches, the overhead remains within a few megabytes (less than 25MB) when dealing with 
 up to 10 neighboring agents.
%
This makes all approaches feasible, although in scenarios with limited communication resources, experience 
 sharing would be the optimal choice, as it provides performance similar to NN averaging but with 
 much lower overhead.

% \begin{itemize}
% \item \textbf{Performance comparison:}
% \begin{itemize}
% \item Present the results of each distributed strategy against the centralized training baseline.
% \item Use graphs and statistical analysis to compare the average episode length and average reward.
% \item Highlight which neighbor-based methods perform comparably to centralized training.
% \end{itemize}
% \item \textbf{Scalability analysis:}
% \begin{itemize}
% \item Describe how the experiments were modified to assess scalability (increasing agents, spawn area).
% \item Show how each method scales with the increasing complexity, using appropriate visualizations.
% \item Discuss the implications of using pre-trained models with a larger number of agents.
% \end{itemize}
% \item \textbf{Communication overhead analysis:}
% \begin{itemize}
% \item Quantify the amount of information exchange for each strategy.
% \item Compare the communication overhead across different neighborhood sizes, possibly using graphs.
% \item Discuss the trade-offs between performance and communication efficiency.
% \item Mention potential techniques for reducing communication overhead.
% \end{itemize}
% \end{itemize}

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{figures/episode_reward_mean.pdf}
  \includegraphics[width=0.8\linewidth]{figures/episode_len_mean.pdf}
  \caption{Data collected during training. 
  We can see that two of the three neighbor-based strategies achieve performance similar to CTDE.
  }
  \label{fig:results}
\end{figure}

\begin{figure*}[htb]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{figures/mean-time-comparison-8-agents.pdf}
      %\caption{Start of the simulation.}
      % \label{fig:zones1}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{figures/mean-time-comparison-16-agents.pdf}
      %\caption{After k time steps of the simulation.}
      % \label{fig:zones2}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{figures/mean-time-comparison-32-agents.pdf}
      %\caption{End of the simulation.}
      % \label{fig:zones3}
  \end{subfigure}
  \caption{ Data collected during evaluation.
  We can see how, increasing the number of agents, the DTDE methods become more unstable and takes more 
  steps to solve the task, while neighbor-based strategies maintain -- on average -- results comparable to CTDE.
  }
  \label{fig:time-eval}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{figures/scalability.pdf}
  \caption{Communication overhead due to different neighboring strategies.
  }
  \label{fig:scalability}
\end{figure}

\section{Conclusion}\label{sec:conclusion}
This paper proposes several neighbor-based strategies to achieve decentralized multi-agent learning (i.e., DTDE ) with performance 
 comparable to centralized approaches (i.e., CTDE). 
%
Specifically, we introduce three approaches, namely: K-NN averaging, K-NN consensus and experience sharing
 which have been evaluated against various baselines, including CTDE, DTDE based on DQN, and MAPPO. 
%
The results show that two of the three proposed methods significantly enhance DTDE performance, 
 achieving outcomes comparable to CTDE, 
 while NN consensus falls short of this goal,
  maintaining performance similar to DTDE. 
%
Furthermore, we analyzed the communication overhead, 
 demonstrating that it remains minimal across all cases, 
 particularly with experience sharing, 
 making it suitable for scenarios with very limited communication resources.

Future works may include several promising directions, namely:
\begin{enumerate*}[label=(\roman*)]
  \item explore the implementation of more advanced consensus algorithms to determine if this strategy can also achieve 
   results comparable to the other two;
  \item develop more sophisticated communication protocols to further reduce communication overhead; and
  \item test these approaches in a wider range of, and more complex, scenarios than the one considered in this paper.
\end{enumerate*}

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{figures/behavior.pdf}
  \caption{Examples of cooperative and uncoordinated behavior. (Left) desired, agents effectively choose different paths to collect different items. (Right) undesired, agents cover the same path to collect the same item.}
  \label{fig:behavior}
\end{figure}

@@

Regarding the proposed neighbor-based approaches, it can be observed that both \emph{experience sharing} and 
 k-NN averaging enhance the performance of DTDE, achieving results comparable to CTDE while also increasing 
 learning stability. 
% 
On the other hand, k-NN consensus maintains performance levels similar to DTDE, which may be attributed to the 
 simplicity of the implemented consensus policy. 
%
Nevertheless, it remains more stable and achieves better performance than MAPPO.

\paragraph{Scalability}
To evaluate the scalability of the learned policies, we conducted several experiments by increasing the number 
 of agents -- specifically to 8, 16, and 32 -- and measuring the time required to complete the task 
 (i.e., $L_mean$).

The results -- \Cref{fig:time-eval} -- show that as the number of agents increases, DTDE exhibits both a higher mean time 
 to solve the task and greater instability. 
%
In contrast, CTDE, k-NN averaging, and experience sharing maintain strong performance. 
%
Similar to the training phase, k-NN consensus does not achieve the same results as the other two neighbor-based 
 approaches but remains on par with DTDE in terms of performance.
\paragraph{Communication overhead}

The proposed neighbor-based approaches, particularly k-NN averaging and experience sharing, achieve performance
 comparable to CTDE while maintaining a decentralized learning structure similar to DTDE. 
% 
However, this entails a communication overhead at runtime due to the information agents must exchange with their neighbors
 -- see \Cref{fig:scalability}. 
%
For k-NN consensus, the overhead remains constant, as each agent selects the best neighbor and only exchanges 
 information with that agent, 
 meaning the amount of data shared does not increase with the number of neighbors. 
% 
In contrast, for k-NN averaging and experience sharing, the overhead scales linearly. 
%
Specifically, the growth in overhead for experience sharing is minimal, as the agents' experiences require very 
 little memory. 
%
For k-NN averaging, however, this increase is more significant, as even a moderately complex neural network 
 has a substantially larger memory footprint.
% 
Nevertheless, for all approaches, the overhead remains within a few megabytes (less than 25MB) when dealing with 
 up to 10 neighboring agents.
%
This makes all approaches feasible, although in scenarios with limited communication resources, experience 
 sharing would be the optimal choice, as it provides performance similar to k-NN averaging but with 
 much lower overhead.
% \begin{itemize}
% \item \textbf{Performance comparison:}
% \begin{itemize}
% \item Present the results of each distributed strategy against the centralized training baseline.
% \item Use graphs and statistical analysis to compare the average episode length and average reward.
% \item Highlight which neighbor-based methods perform comparably to centralized training.
% \end{itemize}
% \item \textbf{Scalability analysis:}
% \begin{itemize}
% \item Describe how the experiments were modified to assess scalability (increasing agents, spawn area).
% \item Show how each method scales with the increasing complexity, using appropriate visualizations.
% \item Discuss the implications of using pre-trained models with a larger number of agents.
% \end{itemize}
% \item \textbf{Communication overhead analysis:}
% \begin{itemize}
% \item Quantify the amount of information exchange for each strategy.
% \item Compare the communication overhead across different neighborhood sizes, possibly using graphs.
% \item Discuss the trade-offs between performance and communication efficiency.
% \item Mention potential techniques for reducing communication overhead.
% \end{itemize}
% \end{itemize}

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{figures/episode_reward_mean.pdf}
  \includegraphics[width=0.8\linewidth]{figures/episode_len_mean.pdf}
  \caption{Training performance of neighbor-based decentralized training strategies compared to centralized (CTDE) and fully decentralized (DTDE) approaches. 
  Two of the three neighbor-based methods achieve performance comparable to CTDE, 
  demonstrating the effectiveness of these methods in achieving scalability and coordination.
  }
  \label{fig:results}
\end{figure}

\begin{figure*}[htb]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{figures/mean-time-comparison-8-agents.pdf}
      %\caption{Start of the simulation.}
      % \label{fig:zones1}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{figures/mean-time-comparison-16-agents.pdf}
      %\caption{After k time steps of the simulation.}
      % \label{fig:zones2}
  \end{subfigure}
  \begin{subfigure}[b]{0.3\textwidth}
      \centering
      \includegraphics[width=\textwidth]{figures/mean-time-comparison-32-agents.pdf}
      %\caption{End of the simulation.}
      % \label{fig:zones3}
  \end{subfigure}
  \caption{Scalability comparison of training methods in the evaluation phase.
  DTDE performance degrades with increasing agent numbers, 
  while CTDE and neighbor-based methods maintain comparable efficiency.
  }
  \label{fig:time-eval}
\end{figure*}

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{figures/scalability.pdf}
  \caption{Communication overhead (in MB) for different neighboring strategies as a function of the number of neighbors.
  }
  \label{fig:scalability}
\end{figure}

\section{Conclusion}\label{sec:conclusion}
This paper proposes several neighbor-based strategies to achieve decentralized multi-agent learning (i.e., DTDE ) with performance 
 comparable to centralized approaches (i.e., CTDE) therefore enabling online adaptation.
%
Specifically, we discuss three approaches, namely: k-NN averaging, k-NN consensus and experience sharing
 which have been evaluated against various baselines, including CTDE, DTDE based on DQN, and MAPPO. 
%
The results show that two of the three proposed methods significantly enhance DTDE performance, 
 achieving outcomes comparable to CTDE, 
 while NN consensus falls short of this goal,
  maintaining performance similar to DTDE. 
%
Furthermore, we analyzed the communication overhead, 
 demonstrating that it remains minimal across all cases, 
 particularly with experience sharing, 
 making it suitable for scenarios with very limited communication resources.

Future works may include several promising directions, namely:
\begin{enumerate*}[label=(\roman*)]
  \item explore the implementation of more advanced consensus algorithms (e.g., deep QD-learning) to determine if this strategy can also achieve 
   results comparable to the other two;
  \item develop more sophisticated communication protocols to further reduce communication overhead; and
  \item test these approaches in a wider range and more complex scenarios than the one considered in this paper (e.g., vectorized multi-agent environment~\cite{bettini2022vmas}).
\end{enumerate*}
